Covering Your Basis: Model Agnostic Probes of the Binary Black Hole Population

### 2022-09-27 src/tex/ms.tex
more refs
Update readme with current tasks / progress
try using caption package -- doesnt work
fix figure icon buttons overlapping fig captions
remove unused latex packages
rework macros and fix some abs things

\documentclass[twocolumn, linenumber]{aastex63}

@@

\usepackage{showyourwork}
\usepackage[perpage]{footmisc}
\usepackage{colortbl}
\usepackage{lineno}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{appendix}
\input{macros}

\definecolor{Gray}{gray}{0.9}
\linenumbers

@@

\begin{abstract}                 
We introduce a non-parametric model built from basis splines, and apply it to infer the astrophysical distribution of 
binary black hole (BBH) masses and spins using the 69 confident observations in GWTC-3. We report the same features previously recovered 
with similarly flexible models in the mass distribution, most notably the peaks in merger rates at \~10\msun and \~35\msun. 
We then use our method as the first non-parametric model applied to the BBH spin distribution, corroborating the results of \citet{spinitasyoulike}, 
showing that tilt distributions may not peak at $\cos(\theta_i) = 1$ as previous models assumed. We find broad agreement with the 
accepted conlcusions of the spin magnitude distribution: the majority of BBH spins are small ($a<0.5$), the distribution peaks at $a\sim0.2$, 
and there is mild support for a non-spinning subpopulaton, but cannot resolve this with current catalog sizes. We also test components whether the data 
prefers the spin distributions to be identical for each component, finding no strong sign of differening distibutions. When modeling the $\chi_\mathrm{eff}$ 
distribution with basis splines we find similar conclusions as previously reported, consistent with a non-isotropic distribution 
(peaks at $\chi_\mathrm{eff}$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][PeakChiEff]}$}), showing considerable support 
($\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][FracBelow0]}$}) for negative effective spins or misaligned systems.
We conclude by discussing how non-parametric methods in gravitatonal-wave population inference are poised to provide a useful complement to 
the usual parametric approach going into the data rich catalog era of gravitational-wave astronomy.
\end{abstract}

@@

\section{Acknowledgements}\label{sec:acknowledments}
We thank Will Farr, Tom Callister, Amanda Farah and Maya Fishbach for useful discussions during the preparation this manuscript, 
along with helpful comments on an early draft by Tom Callister.
This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center 
(\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration. 
This work benefited from access to the University of Oregon high performance computer, Talapas.  This material is based upon work supported 
in part by the National Science Foundation under Grant PHY-1807046 and work supported by NSF’s LIGO Laboratory which is a major facility 
fully funded by the National Science Foundation.
\software{
\textsc{Astropy}~\citep{2018AJ....156..123A},
\textsc{NumPy}~\citep{harris2020array},
\textsc{Matplotlib}~\citep{Hunter:2007},
\textsc{jax},
\textsc{numpyro},
\bruce{CITE ALL ABOVE}
}
\bibliography{bib}{}
\bibliographystyle{aasjournal}

>>>>

\documentclass[twocolumn, linenumbers]{aastex631}

@@

\usepackage{showyourwork}
%\usepackage[perpage]{footmisc}
%\usepackage{colortbl}
%\usepackage{lineno}
%\linenumbers
\usepackage{bm}
\usepackage{amsmath}
\usepackage{appendix}
%\usepackage[margin=10pt]{caption}
\input{macros}

%\definecolor{Gray}{gray}{0.9}

@@

\begin{abstract}                 
We introduce a non-parametric model built from basis splines, and apply it to infer the astrophysical distribution of 
binary black hole (BBH) masses and spins using the 69 confident observations in GWTC-3. We report the same features previously recovered 
with similarly flexible models in the mass distribution, most notably the peaks in merger rates at \~10\msun and \~35\msun. 
We then use our method as the first non-parametric model applied to the BBH spin distribution, corroborating the results of \citet{spinitasyoulike}, 
showing that tilt distributions may not peak at $\cos(\theta_i) = 1$ as previous models assumed. We find broad agreement with the 
accepted conlcusions of the spin magnitude distribution: the majority of BBH spins are small ($a<0.5$), the distribution peaks at $a\sim0.2$, 
and there is mild support for a non-spinning subpopulaton, but cannot resolve this with current catalog sizes. We also test components whether the data 
prefers the spin distributions to be identical for each component, finding no strong sign of differening distibutions. When modeling the $\chi_\mathrm{eff}$ 
distribution with basis splines we find similar conclusions as previously reported, consistent with a non-isotropic distribution 
(peaks at $\chi_\mathrm{eff}=\,$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][PeakChiEff]}$}), showing considerable support 
($f_{\chi_\mathrm{eff}<0}=\,$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][FracBelow0]}$}) for negative effective spins or misaligned systems.
We conclude by discussing how non-parametric methods in gravitatonal-wave population inference are poised to provide a useful complement to 
the usual parametric approach going into the data rich catalog era of gravitational-wave astronomy.
\end{abstract}

@@

\section{Acknowledgements}\label{sec:acknowledments}
We thank Will Farr, Tom Callister, Amanda Farah and Maya Fishbach for useful discussions during the preparation this manuscript, 
along with helpful comments on an early draft by Tom Callister.
This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center 
(\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration. 
The authors are grateful for computational resources provided by the LIGO Laboratory and supported by National Science Foundation Grants PHY-0757058 and PHY-0823459.  
This work benefited from access to the University of Oregon high performance computer, Talapas. This material is based upon work supported 
in part by the National Science Foundation under Grant PHY-1807046 and work supported by NSF’s LIGO Laboratory which is a major facility 
fully funded by the National Science Foundation.
\software{
\textsc{Astropy}~\citep{2018AJ....156..123A},
\textsc{NumPy}~\citep{harris2020array},
\textsc{Matplotlib}~\citep{Hunter:2007},
\textsc{Jax}~\citep{jax},
\textsc{NumPyro}~\citep{pyro,numpyro},
\textsc{GWPopulation}~\citep{Talbot_2019},
\textsc{Bilby}~\citep{Ashton_2019,bilby_gwtc1},
\textsc{Dynesty}~\citep{Speagle_2020}
}
\bibliography{bib}{}
\bibliographystyle{aasjournal}
